---
title: "**Predicting Car Prices**"
author: "*Eric Nguyen*"
date: "*2025-04-27*"
output: html_document
---

### **General Introduction of the Project**
#### Introduction
[...]

#### Purpose
[...]

#### Future Developments
[...]

#### Libraries used
`readr`, `tidyr`, `dplyr`, `tibble`, `tidyverse`, `purrr`, `caret`, `ggplot2``

### **Implementing**
#### Loading libraries
``` {r, message = FALSE, warning = FALSE}
library(tidyr)
library(readr)
library(dplyr)
library(tibble)
library(tidyverse)
library(purrr)
library(caret)
library(ggplot2)
```

#### Bringing in the dataset
``` {r}
cars <- read_csv("C:\\Users\\nguye\\Documents\\Coding\\Personal Projects\\R\\Predicting Car Prices\\data\\imports-85.data")

head(cars, 10)
colnames(cars)
```

As we can see, the column names are wrongly read. Therefore, we'll rename the column names properly.

``` {r}
colnames(cars) = c(
  "symboling",
  "normalized_losses",
  "make",
  "fuel_type",
  "aspiration",
  "num_doors",
  "body_style",
  "drive_wheels",
  "engine_location",
  "wheel_base",
  "length",
  "width",
  "height",
  "curb_weight",
  "engine_type",
  "num_cylinders",
  "engine_size",
  "fuel_system",
  "bore",
  "stroke",
  "compression_ratio",
  "horsepower",
  "peak_rpm",
  "city_mpg",
  "highway_mpg",
  "price"
)

colnames(cars)
head(cars, 10)
```

#### Removing non-numeric columns and missing datas
``` {r}
cars <- cars %>%
  select(
    symboling, 
    wheel_base, 
    length, width, 
    height, 
    curb_weight,
    engine_size, 
    bore, 
    stroke, 
    compression_ratio, 
    horsepower, 
    peak_rpm, 
    city_mpg, 
    highway_mpg, 
    price
  ) %>%
  filter(
    stroke != "?",
    bore != "?",
    horsepower != "?",
    peak_rpm != "?",
    price != "?",
  ) %>%
  mutate(
    stroke = as.numeric(stroke),
    bore = as.numeric(bore),
    horsepower = as.numeric(horsepower),
    peak_rpm = as.numeric(peak_rpm),
    price = as.numeric(price),
  )
```

Double-checking if all the columns are numeric.

``` {r}
map(cars, typeof)
```

#### Examining relationships between predictors
``` {r}
featurePlot(cars, cars$price)
```

There seems to be a fairly-strong positive linear relationship between variable `price` and `horsepower`, `curb_weight`, `engine_size`, `length` and `width`. A slightly-weaker same relationship occurs with the variable `wheel_base`.

On the other hand, variable `city_mpg` and `highway_mpg`have a negative linear relationship with the `price`.

The other variables are likely to have no relationship at all with the `price`.

#### Distribution of the `price` outcome
``` {r, message = FALSE}
ggplot(cars, aes(x = price)) +
  geom_histogram(color = "white") +
  labs(
    title = "Distribution of prices in cars dataset",
    x = "Price",
    y = "Frequency"
  )
```

There seems to be no outliers due to the evenly distribution of the prices in the dataset.

#### Setting up the train-test split
``` {r}
split_indices <- createDataPartition(cars$price, p = 0.8, list = FALSE)
train_cars <- cars[split_indices, ]
test_cars <- cars[-split_indices, ]
```

#### Cross-validation and Hyperarameter Optimization
``` {r}
five_fold_control <- trainControl(method = "cv", number = 5)
tuning_grid <- expand.grid(k = 1:20)
```

#### Experimenting with different models
``` {r}
full_model <- train(price ~ .,
                    data = train_cars,
                    method = "knn",
                    trControl = five_fold_control,
                    tuneGrid = tuning_grid,
                    preProcess = c("center", "scale"))
```

#### Final model evaluation
``` {r}
predictions <- predict(full_model, newdata = test_cars)
postResample(pred = predictions, obs = test_cars$price)
```